---
title: "Classification-as-typical-and-deviant"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Classification-as-typical-and-deviant}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MMRcaseselection)
```

The classification (designation, assignment) of cases as typical and deviant (or another) type needs to be distinguished from the actual choice of cases belonging to a specific type. The package offers two ways of classifying cases. First, one can designate cases using an % prediction interval or the standard deviation of the residuals. Both classification techniques require an `lm` object as input. 

The approach using the prediction interval additionally requires the specification of the width of the interval. The % prediction interval represents the range of outcome values within which 95% of future outcome values are expected to fall. Following Rohlfing and Starke, a case is designated as typical if the observed outcome value is inside the interval and as deviant otherwise. Here is a simple example using the in-built `mtcars` dataset and a 90% predidction interval. 

The output of the function is a dataframe with the observed and fitted outcome value, the upper and lower bound of the prediction interval and the case status based on the fitted value and the prediction interval.
```{r}
df <- lm(mpg ~ disp + wt, data = mtcars)
predint(df, piwidth = 0.9)
```

