---
title: "Pathway case"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pathway-case}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MMRcaseselection)
library(ggplot2) # for illustration of plot customization
```

The pathway case is a type of case that is supposed to make it most likely to find evidence for a mechanism. 
Originally, it was proposed by Gerring ([2007](https://doi.org/10.1177%2F0010414006290784)). The *pathway value* of a case is defined as $|resid_{reduced}-res_{full}|$ if it holds that $|resid_{reduced}| > |resid_{full}|$. In follow-up research, Weller and Barnes ([2014](https://doi.org/10.1017/CBO9781139644501)) propose a different calculation of the pathway value: $|resid_{reduced}|-|res_{full}|$. The package contains a function allowing one to calculate both pathway values and plotting them. 

The function `pathway()` requires the full regression model and the reduced regression model as input. The reduced model omits the pathway variable that is of theoretical interest. Both models must be fed into the fucntion as `lm` objects.
```{r}
df_full <- lm(mpg ~ disp + wt, data = mtcars) # full model
df_reduced <- lm(mpg ~ wt, data = mtcars) # reduced model
pw_out <- pathway(df_full, df_reduced) # calculation of pathway values
head(pw_out)
```

